---
title: "Exercise 1 - Sampling and distributions"
output: html_document
---

```{r loading_libraries, include=FALSE}
library(tidyverse)

# We have provided some additional functions to help with these exercises in the 'stats_helpers.R' file
# We can run that script to make them available here using the `source` function

source('stats_helpers.R')

```

R has a number of inbuilt distributions that are commonly used in statistical analysis.
Sampling from these distributions is done using the `r<distrib>()` functions, where `<distrib>` is 
substituted for the short name of a distribution. Some distributions you might be exposed to 
(and their sampling functions) are:

* The Normal (Gaussian) distribution - `rnorm()`
* The Poisson distribution - `rpois()`
* The Gamma distribution - `rgamma()`
* The Exponential distribution - `rexp()`
* The Binomial distribution - `rbinom()`

Each of these functions are called with an argument `n` for the number of samples to take, as well 
as arguments that specify the parameters of the distribution. For example, a normal distribution is
defined by it's mean and standard deviation, so these can be provided to `rnorm()`

```{r distribution_sampling}
#Using default mean (0) and standard deviation (1)
normal_default <- rnorm(n = 10)

#Setting mean and standard deviation
normal_specified <- rnorm(n = 10, mean = 5, sd = 3)

normal_default

normal_specified
```

Using these sampling functions, complete the following exercises. Some initial code has been 
provided but you will need to fill in the blanks (????? or _____) and extend them to other 
distributions.

>Sample from the Normal and the Exponential distributions.  Take 5, 10, 200 and 1000 samples 
with default parameters and produce a histogram the results.   

```{r changing_sample_size}
# Normal distribution
norm_5 <- rnorm(n = 5)
norm_10 <- rnorm(n = 10)
norm_200 <- rnorm(n = 200)
norm_1000 <- rnorm(n = 1000)

# Histograms for normal distribution
hist(norm_5)
hist(norm_10)
hist(norm_200)
hist(norm_1000)

# Exponential distribution
exp_5 <- rexp(n = 5)
exp_10 <- rexp(n = 10)
exp_200 <- rexp(n = 200)
exp_1000 <- rexp(n = 1000)

# Histograms for exponential distribition
hist(exp_5)
hist(exp_10)
hist(exp_200)
hist(exp_1000)
```

> Take 30 samples and repeat this 6 times for your two distributions. Histogram the observations.
How does the histogram of the samples look different each time?

```{r repeated_sampling}
# Normal distribution
repeated_norm <- tibble(
  a = rnorm(30), 
  b = rnorm(30), 
  c = rnorm(30), 
  d = rnorm(30), 
  e = rnorm(30), 
  f = rnorm(30)
)

repeated_norm

# Exponential distribution
repeated_exp <- tibble(
  a = rexp(n = 30), 
  b = rexp(n = 30), 
  c = rexp(n = 30), 
  d = rexp(n = 30), 
  e = rexp(n = 30), 
  f = rexp(n = 30)
)

# Histograms with hist()
hist(repeated_norm$a)
hist(repeated_norm$b)
hist(repeated_norm$c)
hist(repeated_norm$d)
hist(repeated_norm$e)
hist(repeated_norm$f)

# Histograms with hist()
hist(repeated_exp$a)
hist(repeated_exp$b)
hist(repeated_exp$c)
hist(repeated_exp$d)
hist(repeated_exp$e)
hist(repeated_exp$f)


# Histograms with ggplot and facets
long_norm <- gather(repeated_norm, sample_id, sample_value)

long_norm

long_exp <- gather(repeated_exp, sample_id, sample_value)

long_exp


ggplot(long_norm, aes(x = sample_value)) + geom_histogram() + facet_wrap(~ sample_id)
ggplot(short_norm, aes(x = sample_value)) + geom_histogram(bins = 5) + facet_wrap(~ sample_id)

```

> Create box and whisker plots for your repeated samples. The central box shows between (roughly) 
the quartiles of the data, with the median represented  by a line. The 'whiskers' goes out to the 
extremes of data. Compare with the histograms, do they summarise the distributions well?

By default, the `boxplot()` function will plot one boxplot for each column in your data. This may 
not be suitable if you have a tidy dataframe with one column specifying a grouping variable. In this
case, you can use the formula method to create boxplots of one column based on groupings in a second
column. Both examples are shown below with our wide and long datasets.

Using `ggplot` provides more control over how the boxplots are constructed. But it may be overkill 
if you are just wanting a quick initial look at your data.

```{r boxplots}
# Normal distribution
## Boxplots with boxplot()
boxplot(repeated_norm)

boxplot(sample_value ~ sample_id, data = long_norm)

# Boxplots with ggplot
ggplot(long_norm, aes(x=sample_id, y = sample_value)) + geom_boxplot()

# Exponential distribution
boxplot(repeated_exp)

boxplot(sample_value ~ sample_id, data = long_exp)

# Boxplots with ggplot
ggplot(long_exp, aes(x=sample_id, y = sample_value)) + geom_boxplot()

```

To practice these on more realistic data, we have created several functions that mimic the sort of
data you might encounter in a real situation. These functions all take a single argument `n` that specifies
the number of samples to take:

* `sample_heights(n)`: Sample `n` values from a human height distribution
* `sample_income(n)`: Sample `n` values approximating average weekly income from the 2016 Australian Census
* `sample_nuggety(n)`: Sample `n` values from a "nuggety" distribution, with mostly low value samples,
                       but a few high value "gold nuggets".

>Explore these data sets with histograms and boxplots. What are your initial impressions about 
these data sets as you explore them with different sampling depths?

```{r real_data}
# sample_height with 50 values
hgt50 <- sample_height(50)

repeated_hgt <- tibble(
  a = rnorm(hgt50), 
  b = rnorm(hgt50), 
  c = rnorm(hgt50), 
  d = rnorm(hgt50), 
  e = rnorm(hgt50), 
  f = rnorm(hgt50)
)

repeated_hgt

# Histograms with ggplot and facets
some_hgt50 <- gather(repeated_hgt, sample_id, height50)

some_hgt50

ggplot(some_hgt50, aes(x = sample_id, y = height50)) + geom_boxplot()

ggplot(some_hgt50, aes(x = height50)) + geom_histogram() + facet_wrap(~sample_id)

ggplot(some_hgt50, aes(x = height50)) + geom_histogram(bins = 30) + facet_wrap(~sample_id)


#---------------------------------------------------------------------------------------------#

# sample_height with 1000 values
hgt1000 <- sample_height(1000)

repeated_hgt <- tibble(
  a = rnorm(hgt1000), 
  b = rnorm(hgt1000), 
  c = rnorm(hgt1000), 
  d = rnorm(hgt1000), 
  e = rnorm(hgt1000), 
  f = rnorm(hgt1000)
)

repeated_hgt

# Histograms with ggplot and facets
some_hgt1000 <- gather(repeated_hgt, sample_id, height1000)

some_hgt1000

ggplot(some_hgt1000, aes(x = sample_id, y = height1000)) + geom_boxplot()

#---------------------------------------------------------------------------------------------#

# sample_income with 50 values
inc50 <- sample_income(50)

repeated_inc50 <- tibble(
  a = rnorm(inc50), 
  b = rnorm(inc50), 
  c = rnorm(inc50), 
  d = rnorm(inc50), 
  e = rnorm(inc50), 
  f = rnorm(inc50)
)

repeated_inc50

# Histograms with ggplot and facets
some_inc50 <- gather(repeated_inc50, sample_id, income50)

some_inc50

ggplot(some_inc50, aes(x = sample_id, y = income50)) + geom_boxplot()

#---------------------------------------------------------------------------------------------#
# sample_income with 1000 values
inc1000 <- sample_income(1000)

repeated_inc1000 <- tibble(
  a = rnorm(inc1000), 
  b = rnorm(inc1000), 
  c = rnorm(inc1000), 
  d = rnorm(inc1000), 
  e = rnorm(inc1000), 
  f = rnorm(inc1000)
)

repeated_inc1000

# Histograms with ggplot and facets
some_inc1000 <- gather(repeated_inc1000, sample_id, income1000)

some_inc1000

ggplot(some_inc1000, aes(x = sample_id, y = income1000)) + geom_boxplot()

#---------------------------------------------------------------------------------------------#
# sample_nuggety with 50 values
nug50 <- sample_nuggety(50)

repeated_nug50 <- tibble(
  a = rnorm(nug50), 
  b = rnorm(nug50), 
  c = rnorm(nug50), 
  d = rnorm(nug50), 
  e = rnorm(nug50), 
  f = rnorm(nug50)
)

repeated_nug50

# Histograms with ggplot and facets
some_nug50 <- gather(repeated_nug50, sample_id, nugget50)

some_nug50

ggplot(some_nug50, aes(x = sample_id, y = nugget50)) + geom_boxplot()

#---------------------------------------------------------------------------------------------#
# sample_nuggety with 1000 values
nug1000 <- sample_nuggety(1000)

repeated_nug1000 <- tibble(
  a = rnorm(nug1000), 
  b = rnorm(nug1000), 
  c = rnorm(nug1000), 
  d = rnorm(nug1000), 
  e = rnorm(nug1000), 
  f = rnorm(nug1000)
)

repeated_nug1000

# Histograms with ggplot and facets
some_nug1000 <- gather(repeated_nug1000, sample_id, nugget1000)

some_nug1000

ggplot(some_nug1000, aes(x = sample_id, y = nugget1000)) + geom_boxplot()


```
